---
title: "Final Project"
subtitle: "Affordable Rental Buildings Subsidized by the City of Chicago" 
author: "Daniela Aves"
date: 12/10/24

format: 
  revealjs:
    incremental: true
    theme: serif
    
editor: visual
slide-number: true

---


## Data Sources
My data sources are: 

A list of affordable housing developments that are supported by the City of Chicago to maintain accessible living for Chicagoans 

- https://data.cityofchicago.org/Community-Economic-Development/Affordable-Rental-Housing-Developments/s6ha-ppgi/about_data

The financial statistics for each of the 77 community areas of Chicago

- https://data.cityofchicago.org/Health-Human-Services/Per-Capita-Income/r6ad-wvtk.

## Introduction to Data
In the past, I have explored the availability of youth programs in Chicago through data analysis. In my research, I found that a lot of these programs were free or provided scholarship, and that many programs were funded by the City of Chicago. To further my research into the ways that the City of Chicago creates equity for its residents, I chose a data set containing information on affordable housing funded by the City of Chicago as well as a statistical representation of poverty and hardship levels for each Chicago community area.


```{r}
#| label: load-pkgs-data

# load package(s)
library(tidyverse)
library(ggthemes)
library(patchwork)
library(janitor)
library(lubridate)
library(ggimage)
library(gganimate)
library(ggplot2)
library(dplyr)
library(gt)
library(readr)
library(shiny)

rentals <-
  read_csv("data/Affordable_Rental_Housing.csv")

income <-
  read_csv("data/Per_Capita_Income_Final.csv")

```

## Where do affordable housing units tend to be most common? {.smaller}
- The graph shows the number of affordable rental buildings in each Community Area, the top Community Area being Humboldt Park


```{r}
#| label: table_1

colnames(rentals)[colnames(rentals) == "Community Area Name"] <- "Community_Area_Name"
colnames(income)[colnames(income) == "COMMUNITY AREA NAME"] <- "Community_Area_Name"

unique_com_area <- unique(rentals$Community_Area_Name)

count <- table(rentals$Community_Area_Name)
prc_rentals <- prop.table(count) * 100

table_1 <- data.frame(
  Community_Area_Name = names(count),
  Number_of_Buildings = as.vector(count),
  Percentage = round(as.vector(prc_rentals), 3)
)

ggplot(table_1, aes(x = reorder(Community_Area_Name, Number_of_Buildings),
                    y = Number_of_Buildings)) +
  geom_segment(aes(xend = Community_Area_Name, y = 0, 
                   yend = Number_of_Buildings),
               color = "darkgreen") +
  geom_point(size = 1, color = "darkgreen") +
  labs(
    x = "Community Area Name",
    y = "Number of Affordable Rental Buildings",
    title = "Number of Affordable Rental Builidings per Community Area"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 4.75),
    plot.title = element_text(hjust = 0.5)
  )
```


## Is there a correlation between income per capita and number of affordable housing buildings? {.smaller}

```{r}
#| label: scatter_2

colnames(rentals)[colnames(rentals) == "Community Area Name"] <- "Community_Area_Name"
colnames(income)[colnames(income) == "COMMUNITY AREA NAME"] <- "Community_Area_Name"

unique_com_area <- unique(rentals$Community_Area_Name)

# count is number of buildings
count <- table(rentals$`Community Area Name`)
per_cap_income <- income$`PER CAPITA INCOME`

merged <- merge(rentals, income, by = "Community_Area_Name", all.x = TRUE)

variables <- merged %>%
  group_by(Community_Area_Name) %>%
  summarize(
    val = n(),
    avg_income = first(`PER CAPITA INCOME`)
  )



ggplot(variables, aes(x = val, y = avg_income)) +
  geom_point(color = "darkgreen") +
  geom_smooth(method = "lm", color = "black", se = FALSE) +
  labs(
    x = "Amount of affordable rental buildings",
    y = "Per Capita Income",
    title = "Number of affordable rental bulidings and Per Capita Income"
  ) +
  theme_minimal()
```


## Analysis
We use a line of best fit to see what the correlation is. There is a slight positive correlation, which means that in general as per capita income rises so does the number of affordable rental buildings. This is surprising because an assumption could be made that with a lower average income more affordable housing buildings would be needed for residents. In a more equitable community, we would expect a negative correlation.

## Lastly, what percentage of affordable rental buildings are certain property types?

```{r}
#| label: barchart_3

rentals_prop_type <- rentals %>%
  select(Community_Area_Name, `Property Type`)

results <- rentals_prop_type %>%
  group_by(Community_Area_Name, `Property Type`) %>%
  summarize(count = n(), .groups = 'drop')

ggplot(results, aes(x = Community_Area_Name, y = count, 
                    fill = `Property Type`)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  labs(
    x = "Community Area Name",
    y = "Count of Property Type",
    fill = "Property Type"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 2))

```

## Analysis
The bar charts show percentage of affordable housing buildings that fall into the different property types seen in the legend.

Humboldt Park has the largest number of Multifamily affordable buildings, while West Town has the most Artist related affordable buildings.



